Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-09-27 paper/main.tex
Cut to page limit

Mutation testing, though introduced in the late
1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
has long been a topic of academic interest.  With the recent adoption of mutation
testing by bellwether software industry companies including
Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
Amazon~\cite{AmazonMut}, interest in using mutation testing for
real-world projects has grown widely in practice. 
%
Indeed, interest in mutation testing has become widespread enough in the open
source community that a popular (more than 300 stars on GitHub)
repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
covering almost twenty languages:
\url{https://github.com/theofidry/awesome-mutation-testing}.  Some of
these tools are the products of academic research; others are
essentially the hobby projects of developers interested in enabling
mutation testing for their favorite language.

@@

\begin{table}[hbtp]
\centering
\captionsetup{justification=centering} % Center-align the caption

  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}} \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
\end{table}

@@

%\begin{tabular}{l|r|rr|rr}
% ---------------- C++/Java/Rust table UM Only--------------------------
\begin{table}[hbtp]
\centering
%\captionsetup{justification=centering}
\footnotesize
\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile. Mut. Score refers to the Mutation Score.}
\label{tab:table_cpp1javarust}

@@

% ---------------- C++ table (UM vs Others)--------------------------
\begin{table}[htbp]
\centering
%\small
\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool.}
\label{tab:table_cpp2}
\begin{varwidth}{.8\columnwidth}
\begin{tabularx}{\linewidth}{X|rrr|rr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
 ompl &  564 & 0.07 & 26.7\% & 278 & 0.08  \\
 libgraphqlparser  &  131 & 0.90 & 33.3\% & 52 & 0.96  \\
 rethinkdb\_rebirth &  492 & 0.10 & 12.7\% & 25 & 0.36 \\
 xoreos &  2280 & 0.01 & 3.3\% & 414 & 0.02  \\
 libxmljs &  790 & 0.64 & 6.7\% & 830 & 0.46 \\[0.5ex]
 tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
\textbf{Mean}   & 837.83 & 0.29 & 13.67\% & 276.67 & 0.33\\
\textbf{Std.Dev.} & 745.78 & 0.37 & 13.50\%   & 311.56 & 0.38 \\\bottomrule
\end{tabularx}
\end{varwidth}
\end{table}

@@

\begin{table}[htbp]
\centering
%\small
\captionsetup{justification=centering}
\caption{\small Java (Our Implementation vs. PIT vs. LittleDarwin).
  LittleDarwin and PIT produce no equivalent mutants.}
\label{tab:table_java2}

@@

\begin{table}[htbp]
\centering
\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\small
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 6.7\% & 482 & 0.72 & 20.0\% & 921 & 0.27 & 3.3\% \\
dtw & 866 & 0.55 & 0\% & 270 & 0.44 & 3.3\% & 713 & 0.46 & 27.7\% \\
notion-sdk-py & 196 & 0.60 & 13.3\% & 63 & 0.65 & 0\% & 52 & 0.44 & 13.9\% \\
atlassian- & \mr{2}{333} & \mr{2}{0.62} & \mr{2}{30.0\%} & \mr{2}{174} & \mr{2}{0.71}  & \mr{2}{13.3\%} & \mr{2}{106} & \mr{2}{0.74} & \mr{2}{3.5\%}  \\
python-api  &    &    &    &    &   &   &   &  &  \\
ESD & 1712 & 0.07 & 6.7\% & 720 & 0.08 & 13.3\% & 1013 & 0.06 & 6.7\% \\
Fiber & 458 & 0.64 & 23.3\% & 222 & 0.57 & 46.7\% & 320 & 0.81 & 20.0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 13.3\% & 321.83 & 0.53 & 16.1\% & 520.93 & 0.46 & 12.34\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.4\% & 239.04 & 0.24 & 16.7\% & 417.50 & 0.28 & 9.57\% \\\bottomrule        
\end{tabular}
\end{table}

@@

\begin{table}[htbp]
\centering
%\small
\caption{Rust (Our Implementation vs. Cargo-Mutants)}
\label{tab:table_rust2}
\begin{varwidth}{.8\linewidth}
\begin{tabularx}{\columnwidth}{X|rrr|rrr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{3}{c}{\textbf{Cargo-Mutants}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\\midrule
cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
passarine & 173 & 0.52 & 26.7\% & 0 & 0.00 & 0\%   \\
typos & 366 & 0.77 & 10.0\% & 6 & 0.83 & 0\%   \\
ord & 643 & 0.83 & 20.0\% & 6 & 1.00 & 0\%   \\
bazuka & 402 & 0.73 & 16.7\% & 0 & 0.00 & 0\%   \\
strum & 50 & 0.22 & 6.7\% & 2 & 1.00 & 0\%   \\\midrule
\textbf{Mean}   & 554.00 & 0.64 & 13.3\% & 3.17 & 0.54 & 0.00\%\\
\textbf{Std. Dev} & 592.63 & 0.25 & 9.7\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
\end{tabularx}
\end{varwidth}
\end{table}

@@

As future work, we plan to give users more control over which mutation
operators are used, with controls common to multiple languages.
Our approach in general seems to be a good fit to modern
large software projects, which are often
multilingual~\cite{Multilingual}: a single mutation tool can be used
for an entire project.

>>>>

Mutation testing, though introduced in the late
1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
has long been a topic of academic interest.  With the recent adoption of mutation
testing by bellwether software industry companies including
Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
Amazon~\cite{AmazonMut}, interest in using mutation testing for
real-world projects has grown widely in practice. 
%
Indeed, interest in mutation testing has become widespread enough in the open
source community that a popular (more than 300 stars on GitHub)
repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
covering almost twenty languages.\footnote{\url{https://github.com/theofidry/awesome-mutation-testing}}  Some of
these tools are the products of academic research; others are
essentially the hobby projects of developers interested in enabling
mutation testing for their favorite language.

@@

\begin{table}[hbtp]
\centering
\small
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}} \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\label{tab:rules}
\end{table}

@@

%\begin{tabular}{l|r|rr|rr}
% ---------------- C++/Java/Rust table UM Only--------------------------
\begin{table}[hbtp]
\centering
\small
%\captionsetup{justification=centering}
\footnotesize
\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile. Mut. Score refers to the Mutation Score.}
\label{tab:table_cpp1javarust}

@@

% ---------------- C++ table (UM vs Others)--------------------------
\begin{table}[htbp]
\centering
\small
\caption{C++ (Comby vs. Dextool). We cannot measure equivalent mutants for Dextool.}
\label{tab:table_cpp2}
\begin{varwidth}{.8\columnwidth}
\begin{tabularx}{\linewidth}{X|rrr|rr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
 ompl &  564 & 0.07 & 26.7\% & 278 & 0.08  \\
 libgraphqlparser  &  131 & 0.90 & 33.3\% & 52 & 0.96  \\
 rethinkdb\_rebirth &  492 & 0.10 & 12.7\% & 25 & 0.36 \\
 xoreos &  2280 & 0.01 & 3.3\% & 414 & 0.02  \\
 libxmljs &  790 & 0.64 & 6.7\% & 830 & 0.46 \\[0.5ex]
 tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
\textbf{Mean}   & 837.83 & 0.29 & 13.67\% & 276.67 & 0.33\\
\textbf{Std.Dev.} & 745.78 & 0.37 & 13.50\%   & 311.56 & 0.38 \\\bottomrule
\end{tabularx}
\end{varwidth}
\end{table}

@@

\begin{table}[htbp]
\centering
\small
\caption{Java (Comby vs. PIT vs. LittleDarwin).
  LittleDarwin and PIT produce no equivalent mutants.}
\label{tab:table_java2}

@@

\begin{table}[htbp]
\centering
\caption{Python (Comby vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\small
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 6.7\% & 482 & 0.72 & 20.0\% & 921 & 0.27 & 3.3\% \\
dtw & 866 & 0.55 & 0\% & 270 & 0.44 & 3.3\% & 713 & 0.46 & 27.7\% \\
notion-sdk-py & 196 & 0.60 & 13.3\% & 63 & 0.65 & 0\% & 52 & 0.44 & 13.9\% \\
atlassian- & \mr{2}{333} & \mr{2}{0.62} & \mr{2}{30.0\%} & \mr{2}{174} & \mr{2}{0.71}  & \mr{2}{13.3\%} & \mr{2}{106} & \mr{2}{0.74} & \mr{2}{3.5\%}  \\
python-api  &    &    &    &    &   &   &   &  &  \\
ESD & 1712 & 0.07 & 6.7\% & 720 & 0.08 & 13.3\% & 1013 & 0.06 & 6.7\% \\
Fiber & 458 & 0.64 & 23.3\% & 222 & 0.57 & 46.7\% & 320 & 0.81 & 20.0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 13.3\% & 321.83 & 0.53 & 16.1\% & 520.93 & 0.46 & 12.34\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.4\% & 239.04 & 0.24 & 16.7\% & 417.50 & 0.28 & 9.57\% \\\bottomrule        
\end{tabular}
\end{table}

@@

\begin{table}[htbp]
\centering
\small
\caption{Rust (Comby vs. Cargo-Mutants)}
\label{tab:table_rust2}
\begin{varwidth}{.8\linewidth}
\begin{tabularx}{\columnwidth}{X|rrr|rrr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{3}{c}{\textbf{Cargo-Mutants}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\\midrule
cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
passarine & 173 & 0.52 & 26.7\% & 0 & 0.00 & 0\%   \\
typos & 366 & 0.77 & 10.0\% & 6 & 0.83 & 0\%   \\
ord & 643 & 0.83 & 20.0\% & 6 & 1.00 & 0\%   \\
bazuka & 402 & 0.73 & 16.7\% & 0 & 0.00 & 0\%   \\
strum & 50 & 0.22 & 6.7\% & 2 & 1.00 & 0\%   \\\midrule
\textbf{Mean}   & 554.00 & 0.64 & 13.3\% & 3.17 & 0.54 & 0.00\%\\
\textbf{Std. Dev} & 592.63 & 0.25 & 9.7\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
\end{tabularx}
\end{varwidth}
\end{table}

@@

As future work, we plan to give users more control over which mutation
operators are used, with controls common to multiple languages.
Our approach is a good fit to modern
large software projects, which are often
multilingual~\cite{Multilingual}: a single mutation tool can be used
for an entire project.
